---
title: "Quality of Weight Lifting Exercises Recognition"
author: "Alexandre Rodichevski"
date: "2015-01-18"
output: html_document
---

##Abstract

This report is an excercise variation of the [research work](http://groupware.les.inf.puc-rio.br/har). We build a machine learning algorithm to predict activity quality from activity monitors.

```{r libraries, echo=FALSE, results='hide'}
library(caret)
```

##Dataset

We download the training dataset.

```{r read-dataset, cache=TRUE}
url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
fn <- "pml-training.csv"
download.file(url, fn, "curl")
pml <- read.csv(fn)
```

We select the only useful columns: **classe** (the activity quality) and ones whose name begins with **roll**, **pitch**, **yaw**, **total**, **gyros**, **accel** and **magnet** (the accelerometer data). We have not considered the columns containing mostly not avaiable values.

```{r subset, cache=TRUE}
pml1 <- pml[grepl("^classe|^roll|^pitch|^yaw|^total|^gyros|^accel|^magnet", names(pml))]
```

##Machine Learning Algorithm

We have used random forest algorythm proposed in the original research.

```{r fitting, cache=TRUE}
modFit <- train(classe ~ ., method = "rf", data = pml1)
```

##Out-of-sample Error

The accuracy is about 99%. The out-of-bag error of the random forest algorythm is about 0.5%. The out-of-sample error is expected to have similar value, about 0.5%.

```{r fitting-view, cache=TRUE}
modFit
model <- modFit$finalModel
median(model$err.rate[, "OOB"])
```

The simple decision tree algorytm is much less accurate, giving about 50% of accuracy.

```{r read-test, cache=TRUE, echo=FALSE, results='hide'}
url2 <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
fn2 <- "pml-testing.csv"
download.file(url2, fn2, "curl")
test <- read.csv(fn2)
test1 <- test[grepl("^classe|^roll|^pitch|^yaw|^total|^gyros|^accel|^magnet", names(test))]
```

```{r predict, cache=TRUE, echo=FALSE, results='hide'}
answers = predict(modFit, test1)
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
pml_write_files(answers)
```

